---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# CPAT

**CPAT** is a package implementing some statistical tests for detecting
structural change in a series of data. The tests made publicly available are:

* The CUSUM test, via `CUSUM.test()`
* The Darling-Erdös test, via `DE.test()`
* The Hidalgo-Seo test, via `HS.test()`
* Andrews test, via `Andrews.test()`
* The Rényi-type test via `HR.test()`

This package was written to facilitate the simulations performed in
a paper by Horváth, Rice and Miller (see the documentation for `HR.test()` for a
citation) and thus is geared to change point tests capable of detecting
early/late changes in a sample. That said, it is general purpose.

## Change Point Testing

Change point testing is performed on sequential data (such as time series) to
determine whether the data shares a common structure. In particular, let $X_t =
\mu_t + \epsilon_t$ where $\epsilon_t$ is a random noise process, with $1 \leq t
\leq T$. With the exception of Andrews' test, the tests mentioned above can be
used to decide between the hypotheses:

$$H_0 : \mu_1 = \ldots \mu_T = \mu$$

$$H_A : \mu_1 = \ldots \mu_{t^{*} - 1} \neq \mu_{t^{*}} = \ldots = \mu_T$$

$t^{*}$ is an integer satisfying $1 \leq t^{*} \leq T$. Critically, $t^{*}$ is
not assumed to be known, so the alternative hypothesis states that the change
occurs at an unknown location in the sample.

Andrews' test is an exception; his test assumes some information about where the
change occured. Suppose $\mu_t = \mu$ for $1 \leq t \leq T^{'}$ with $T^{'} < T$
and $T^{'}$ known. His test decides between the hypotheses:

$$H_0: \mu_{T^{'} + 1} = \ldots = \mu_{T}$$

$$H_A: \mu_{T^{'} + 1} = \ldots = \mu_{t^{*} - 1} \neq \mu_{t^{*}} = \ldots =
\mu_{T}$$

In this case, $T^{'} \leq t^{*} \leq T$.

Change point testing traces its roots to quality control procedures; if one
imagines $X_t$ being some measurement of a part produced by an machine, $H_0$
states that the machine was always calibrated while the alternative hypothesis
claimes that the machine became uncalibrated at some unknown porint in time.
Another view of a change point test is that it's yet another test to check if a
time series is stationary, which is a critical assumption made in most analyses
involving time series.

While the above formulation describes a test primarily concerned about the mean
of a process, these tests, even in this form, can make statements about
structural change other than the mean. For instance, if $X_t$ represents the
residuals of a regression model, the test checks whether the parameters of the
model are stable over time or not.

All of the tests included in this package are asymptotic tests; the test
performs better for large $T$ and one should be cautious when using these tests
for small $T$. Simulations studies suggest that the Rényi-type test seems to
perform best when a change occurs near the ends of a sample, while the CUSUM
test seemed to perform best when the change occured mid-sample.

## Example

Each function can accept a univariate dataset as input, and will return an
`htest`-class object with the results of the test.

```{r example}
library(CPAT)

set.seed(20180924)
x <- c(rnorm(250), rnorm(50) + 1)

CUSUM.test(x)
DE.test(x)
HS.test(x)
HR.test(x)
Andrews.test(x, M = 250)
```

Often these tests return a test statistic, a $p$-value, and the estimated
location of the change. This latter quantity is the arg-max of the terms the
test statistics maximize. (There is theory supporting this estimator for the
CUSUM test, but no theory for the Rényi-type test.)

`Andrews.test()` also allows for testing for structural change in a linear model
directly.

```{r}
df <- data.frame(x = x, y = 1 + 2 * x + c(rnorm(250), rnorm(50) + 1))
Andrews.test(df, M = 250, formula = y ~ x)
```
